Relationships in the extremes and their influence on competition and coexistence

### 2021-09-08 paper/Paper.Rnw
added refs

%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
%biodiversity (Chesson, 1994).
%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
%(more than the abstracts if you want), and consider using them as references here.
The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
the other and drive it extinction \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how stable species coexistence 
nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}, 
%***DAN: 
%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
%right column on the first page of the Barabas review (Ecol Monogr, 2018).
even though it is mathematically complex in the eyes
of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
%more detail than you previously gave to do that.

@@

Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA). Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see Ghosh for impact of ATA on community aggregate properties).

>>>>

%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
%biodiversity (Chesson, 1994).
%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
%(more than the abstracts if you want), and consider using them as references here.
The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
the other and drive it extinction \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how stable species coexistence 
nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}, 
%***DAN: 
%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
%right column on the first page of the Barabas review (Ecol Monogr, 2018).
even though it is mathematically complex in the eyes
of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
%more detail than you previously gave to do that.

@@

Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA). Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
